A weakly structured stem for human origins in Africa

### 2022-07-29 paper.tex

Evidence for archaic hominin admixture in Eurasia has bolstered the plausibility of 
archaic hominin admixture having also occurred in Africa. For this reason, previous 
work has focused on archaic hominin admixture to explain patterns of polymorphism 
inconsistent with a single-origin model. Here, we have shown that weakly-structured-stem 
models better capture these patterns. Preferring weakly-structured-stem models over 
archaic-admixture models potentially has ecological implications. With a weakly 
structured stem, there is no need to posit that an archaic hominin population stayed 
reproductively isolated from the ancestral human lineage for thousands of years before 
the initiation of gene flow. Instead, there would have been continuous or recurrent 
contact between two or more groups present in Africa. It would be fruitful to 
investigate the ecological reasons for continued contact between groups within Africa 
but long periods of isolation between African and Eurasian groups.

>>>>

Evidence for archaic hominin admixture in Eurasia has bolstered the plausibility of 
archaic hominin admixture having also occurred in Africa. For this reason, previous 
work has focused on archaic hominin admixture to explain patterns of polymorphism 
inconsistent with a single-origin model. Here, we have shown that weakly-structured-stem 
models better capture these patterns. Preferring weakly-structured-stem models over 
archaic-admixture models potentially has ecological implications. With a weakly 
structured stem, there is no need to posit that an archaic hominin population in Africa stayed 
reproductively isolated from the ancestral human lineage for tens thousands of years before 
the initiation of gene flow. Instead, there would simply have been continuous or recurrent 
contact between two or more groups present in Africa.
